---
title: "SLR Diamonds"
author: "John Leraas"
date: "10/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Data
data: full data set
data_r: reduced data set including only y values below a certain threshold (threshold part of exploration)

```{r}
data <- read.csv("diamonds4.csv", header=TRUE)
attach(data)
names(data)
summary(data)

```

## Simple Linear Regression

Fit simple linear regression model with no transformatio

```{r }
results <- lm(price~carat)
summary(results)

plot(carat, price, xlab="carat", ylab = "Price $", main="Price vs. Carat")
abline(results, col="red")
```

Clearly a nonlinear relationship exists. Explore exponential relationship: y^lambda = B0 + x*B1

## Perform Box Cox Method

```{r }
library(MASS)
boxcox(results)
boxcox(results, lambda = seq(-0, .5, .005))
```
Assume lambda = 0.333 given plot

## Perform appropriate transformation and replot


```{r}
price.t = price**(.3)
results.t <- lm(price.t~carat)


summary(results.t)

plot(carat, price.t, xlab="carat", ylab = "Price^(1/3) $", main="Price^(.3) vs. Carat")
abline(results.t, col="red")

```

## Replot Box-Cox to test

```{r}
boxcox(results.t)
boxcox(results.t, lambda = seq(-1, 1.5, .01))

```
NOTE: I INIITIALLY USED THE VALUE LAMBDA = 1/3; HOWEVER WHEN I REFIT THE BOX COX PLOT, LAMBDA = 1 FELL OUTSIDE TEH 95% CONFIDENCE INTERVAL

Settled on lambda = 0.3

## Test linear regression assumptions:
1) A linear relationship exists between response & predictor variable (done - see preceding scatterplot)
2) The error terms have mean 0 (residual plot)
3) The error terms have constant variance (residual plot)
4) The error terms are uncorrelated (ACF Plot)
5) The errors folow a normal distribution (QQ Plot)

To start, create Residual Plot. If variance not constant, transform response (y). If mean of error terms not zero, transform predictor variable (x)

```{r}
plot(results.t$fitted.values, results.t$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")
```
Note: this still requires a y-transformation given that the variance does not appear constant (variance gets bigger as price increased)



## Try another model: e^x (also tried 2^x, 10^x with similar results)

```{r}
price.log = log(price)
results.log <- lm(price.log~carat)


summary(results.log)

plot(carat, price.log, xlab="carat", ylab = "ln(Price) $", main="ln(Price) $ vs. Carat")
abline(results.log, col="red")
```


###################################################################################
## STOP AND EVALUATE:
y^0.3 = mx+b seems to be the best fit so far, however this does not meet normality assumptions.
--> Try working with 'reduced' data set
###################################################################################

## Evaluate "Reduced" Data Set

Start by plotting
```{r}
data_r <- data[which(data$price < 30000), ]
summary(data_r)

plot(carat, price, xlab="carat", ylab = "Price $", main="Price vs. Carat")
points(data_r$carat, data_r$price, pch=2, col="red")

```

## Simple Linear Regression - Reduced dataset
Perform simple linear regression on the reduced data set and plot

```{r}
detach(data)
attach(data_r)
results_r <- lm(price~carat)

plot(carat, price, xlab="carat", ylab = "Price $", main="Price vs. Carat")
abline(results_r, col="red")
```

Doesn't look linear, but plot Residual Plot for additional detail

```{r}
plot(results_r$fitted.values, results_r$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")
```
Variance not constant -> need to perform y transformation

Looking at scatterplot, equation looks like an exponential y = c*x^l 
Try Box Cox again... method
1) Make Box Cox Plot
2) Apply transformation
3) Replot to check


## Perform Box Cox Method

```{r }
boxcox(results_r)
boxcox(results_r, lambda = seq(0, .5, .005))
```
Assume lambda = 0.275 given plot

## Perform appropriate transformation and replot


```{r}
price.rt = price**(.275)
results.rt <- lm(price.rt~carat)


summary(results.rt)

plot(carat, price.rt, xlab="carat", ylab = "Price^(0.15) $", main="Price^(0.15) vs. Carat - REDUCED")
abline(results.rt, col="red")

```
Looks ok, though larger carat values don't seem to fit plot as well


## Replot Box-Cox to test

```{r}
boxcox(results.rt)
boxcox(results.rt, lambda = seq(-1, 2, .01))

```
Lambda = 1 falls within 95% confidence

## Residual Plot

```{r}
plot(results.rt$fitted.values, results.rt$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")
```

## Try another model: e^x (also tried 2^x, 10^x with similar results)

```{r}
price_r.log = log(price)
results_r.log <- lm(price_r.log~carat)


summary(results_r.log)

plot(carat, price_r.log, xlab="carat", ylab = "ln(Price) $", main="ln(Price) $ vs. Carat")
abline(results_r.log, col="red")
```
Doesn't look great, but try plotting residuals


## Residual Plot

```{r}
plot(results_r.log$fitted.values, results_r.log$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")
```
###################################################################################
## STOP AND EVALUATE:
Normality assumptions not met by y^lambda -> y' (x^1/lambda) or ln(y) -> y' (e^x) transformations
This is true of full data set and 'reduced' data set
I tried different ranges within the reduced data set (Price < 30k, 10k, 8k, 5k)
-- The same trends persist. 
---- y^lambda seems to provide the best fit, but right side (higher carats) actual values fall below predicted
---- Never meet normality assumptions

###################################################################################

Follow Xin's suggestion: ln(Price) vs. ln(Carat)

```{r}
detach(data_r)
attach(data)

price.log = log(price)
carat.log = log(carat)
results.loglog <- lm(price.log~carat.log)

summary(results.loglog)

plot(carat.log, price.log, xlab="ln(carat)", ylab = "ln(Price) $", main="ln(Price) $ vs. ln(Carat)")
abline(results.loglog, col="red")
```

Plot looks good. Check linearity assumptions
- Residual Plot
- ACF Plot
- QQ Plot

```{r}
#Residual Plot
plot(results.loglog$fitted.values, results.loglog$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")

#ACF Plot
acf(results.loglog$residuals, main="ACF of Residuals")

#QQ Plot
qqnorm(results.loglog$residuals)
qqline(results.loglog$residuals, col="red")
```
Residual Plot - a little unclear if mean = 0 for smallest values; variance of residuals potentially constant
ACF Plot - not quite met (lag = 4, potentially 19)
QQ - not met, but model can still be robust
