---
title: "SLR Diamonds"
author: "John Leraas"
date: "10/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
```

## Import Data
data: full data set
data_r: reduced data set including only y values below a certain threshold (threshold part of exploration)

```{r}
data <- read.csv("diamonds4.csv", header=TRUE)
attach(data)
names(data)
summary(data)

```


## Test linear regression assumptions:
1) A linear relationship exists between response & predictor variable
2) The error terms have mean 0 (residual plot)
3) The error terms have constant variance (residual plot)
4) The error terms are uncorrelated (ACF Plot)
5) The errors follow a normal distribution (QQ Plot)

If variance not constant, transform response (y). If mean of error terms not zero, transform predictor variable (x)





## Plot data 

Fit simple linear regression model with no transformation and plot the data

```{r }
results <- lm(price~carat)

plot(carat, price, xlab="carat", ylab = "Price $", main="Price vs. Carat")
abline(results, col="red")

plot(results$fitted.values, results$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")
```

Clearly a nonlinear relationship exists. Given that variance is not constant, start by ransforming y

## Y Transformation
Explore exponential relationship: y^lambda = B0 + x*B1.
After exploring various values of lambda, we chose lambda = 0 resulting in the transformation y' = ln(y)

Plot data and residual plot:

```{r}
price.log = log(price)
results.log <- lm(price.log~carat)

#Scatter Plot
plot(carat, price.log, xlab="carat", ylab = "ln(Price) $", main="ln(Price) $ vs. Carat")
abline(results.log, col="red")

#Residual Plot
plot(results.log$fitted.values, results.log$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")
```

Note: variance appears to be constant, though the mean is not zero so transform x.



## x Transformation

After trying different possibilities, we chose x' = ln(x)

Plot data, residuals, ACF, and QQ

```{r}
carat.log = log(carat)
results.loglog <- lm(price.log~carat.log)

#Scatter Plot
plot(carat.log, price.log, xlab="ln(carat)", ylab = "ln(Price) $", main="ln(Price) $ vs. ln(Carat)")
abline(results.loglog, col="red")

#Residual Plot
plot(results.loglog$fitted.values, results.loglog$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")

#ACF Plot
acf(results.loglog$residuals, main="ACF of Residuals")

#QQ Plot
qqnorm(results.loglog$residuals)
qqline(results.loglog$residuals, col="red")

#Summary
summary(results.loglog)
anova(results.loglog)


```
Observations:
-- Scatterplot & Residual Plots - look good. Some potential issues for small carat values, but overall good fit
-- ACF - Doesn't quite meet assumption (see lag = 4), however overall very close
-- QQ Plot - residuals do not fit a normal distribution, however model may still be robust



## Box Cox check

Note: while applying Box Cox to the above equation does not result in lambda = 1, it still fits the data reasonably well. Potentially explore easing signficance level.

```{r}
boxcox(results.loglog)
boxcox(results, lambda = seq(-1, 2, .05))
```












