---
title: "STAT 6021 - Diamonds MLR Work"
author: "Arne Newman"
date: "10/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Data
data: full data set
data_r: reduced data set including only y values below a certain threshold (threshold part of exploration)

```{r}
data <- read.csv("diamonds4.csv", header=TRUE)
attach(data)
names(data)
summary(data)

# Create a reduced subset of price <= $30,000
data_r <- data[which(data$price <= 30000), ]

```
### Work on Clarity as Indicator Variable

Check whether Clarity is being treated as a factor.

```{r}
# Check if it is being treated as numeric
is.numeric(clarity) 
##have R treat mfr as categorical
clarity<-factor(clarity) 
is.factor(clarity) 
##check coding scheme
contrasts(clarity)
```
## FLAG - CONSIDER GROUPING 

### Full Data Basic MLR with Carats and Clarity, No Interaction
First run multiple linear regressions (MLRs) for the full and reduced dataset using 
carats and clarity, without interaction. 
```{r}
mlr_clarity <- lm(price~carat+clarity)
summary(mlr_clarity)
```
Let's look at our regression assumptions. 

```{r}
#Residual Plot
plot(mlr_clarity$fitted.values, mlr_clarity$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")

#ACF Plot
acf(mlr_clarity$residuals, main="ACF of Residuals")

#QQ Plot
qqnorm(mlr_clarity$residuals)
qqline(mlr_clarity$residuals, col="red")
```
Residuals have a clear pattern with an increasing variance, the ACH plot doesn't pass the test, and the QQ plot is problematic. 

### Reduced Data Basic MLR with Carats and Clarity, No Interaction
Now let's try this on the reduced dataset, as it may be driven by outliers. 

```{r}
mlr_clarity_red <- lm(price~carat+clarity, data = data_r)
summary(mlr_clarity_red)
```

The $R^2$ for the reduced set is much higher, at `r mlr_clarity_red$r.squared` compared to 
`r mlr_clarity$r.squared` for the full dataset. Let's look at out regression assumptions 
for the reduced dataset. 

```{r}
#Residual Plot
plot(mlr_clarity_red$fitted.values, mlr_clarity_red$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")

#ACF Plot
acf(mlr_clarity_red$residuals, main="ACF of Residuals")

#QQ Plot
qqnorm(mlr_clarity_red$residuals)
qqline(mlr_clarity_red$residuals, col="red")
```
This is better, but our assumptions are all still invalid based on the plots.Let's look for interaction
Break the data into subsets for plotting
```{r}
##consider each region a subset
r1<-subset(data,clarity=="IF") 
r2<-subset(data,clarity=="SI1") 
r3<-subset(data,clarity =="SI2") 
r4<-subset(data,clarity=="VS1") 
r5<-subset(data,clarity=="VS2") 
r6<-subset(data,clarity =="VVS1")
r7<-subset(data,clarity=="VVS2") 

```

Generate regression lines for each race
```{r}
##fit 3 separate regressions, one for each region
reg1<-lm(price~carat,data=r1)
reg2<-lm(price~carat,data=r2)
reg3<-lm(price~carat,data=r3)
reg4<-lm(price~carat,data=r4)
reg5<-lm(price~carat,data=r5)
reg6<-lm(price~carat,data=r6)
reg7<-lm(price~carat,data=r7)
```

Create a scatterplot to view the data, with different symbols and colors for each race. 
```{r}
plot(carat, price, main="Price against carats, by Clarity")
points(r1$carat,r1$price, pch=1, col="black") 
points(r2$carat,r2$price, pch=2, col="red") 
points(r3$carat,r3$price, pch=3, col="blue")
points(r4$carat,r4$price, pch=4, col="green") 
points(r5$carat,r5$price, pch=5, col="yellow")
points(r6$carat,r6$price, pch=6, col="pink") 
points(r7$carat,r7$price, pch=7, col="orange")
abline(reg1,lty=1)
abline(reg2,lty=2, col="red") 
abline(reg3,lty=3, col="blue")
abline(reg4,lty=1, col = "green")
abline(reg5,lty=2, col="yellow") 
abline(reg6,lty=3, col="pink")
abline(reg7,lty=4, col="orange")
legend("topleft", c('IF', 'SI1', 'SI2', 'VS1', 'VS2', 'VVS1', 'VVS2'), lty=c(1,2,3,4,5,6,7), pch=c(1,2,3,4,5,6,7), col=c("black","red","blue", "green", "yellow","pink","orange")) 
```

These slopes area all different, indicating an interaction effect. We thought that grouping the 
clarity ratings with the same letters (e.g., VS1, VS2) into groups may be useful, but this
graph does not visually support that grouping. 


### Reduced Data Basic MLR with Carats and Clarity, No Interaction
Now let's test clarity and cut on the reduced dataset, with interaction. 

```{r}
mlr_clarity_red_inter <- lm(price~carat*clarity, data = data_r)
summary(mlr_clarity_red_inter)
```
The $R^2$ has increased compared to the model without interaction, at `r mlr_clarity_red_inter$r.squared` 
compared to `r mlr_clarity_red$r.squared` for the reduced dataset without interaction. 

Now let's revisit our regression assumptions for this model. 

```{r}
#Residual Plot
plot(mlr_clarity_red_inter$fitted.values, mlr_clarity_red_inter$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")

#ACF Plot
acf(mlr_clarity_red_inter$residuals, main="ACF of Residuals")

#QQ Plot
qqnorm(mlr_clarity_red_inter$residuals)
qqline(mlr_clarity_red_inter$residuals, col="red")
```
We see that mean error is not zero and variance is nonconstant. Let's try a transformation, starting
by transforming the response to address the variance. With carat as a major driver, let's start 
by testing the same transformation as worked for SLR, a log-log transformation - where we use the log of
the response and carat, while leaving the indicator variables. 

```{r}
price_log <- log(price)
carat_log <- log(carat)

mlr_clarity_red_inter_log <- lm(price_log~carat_log*clarity)
summary(mlr_clarity_red_inter_log)
```
Here we see an extremely high $R^2$, but the clarity terms do not look meaningful.
Let's start by reviewing our regression assumptions.


```{r}
#Residual Plot
plot(mlr_clarity_red_inter_log$fitted.values, mlr_clarity_red_inter_log$residuals, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Vaues")
abline(h=0, col="red")

#ACF Plot
acf(mlr_clarity_red_inter_log$residuals, main="ACF of Residuals")

#QQ Plot
qqnorm(mlr_clarity_red_inter_log$residuals)
qqline(mlr_clarity_red_inter_log$residuals, col="red")
```

Our assumptions generally look good here... expand more. 

## Test removal of interaction terms 

